### 2024-09-13 parts/appendix.tex
Hilbert!

\begin{proofof}[undecidable-non-negative:lem]
    Recall that by the undecidability of the Hilbert's Tenth Problem \cite[Problem 10 page 22]{HILB1902}, it is
    undecidable whether a given polynomial $P \in \Rel[\vec{X}]$ has a solution in
    $\Rel^n$ \cite{MATI1970,DAVIS1973}.
    The result follows by observing that
    checking whether a polynomial $P \in \Rel[\vec{X}]$ has no solutions in $\Rel^n$ if and only 
    if $P^2 - 1$ is \kl{non-negative}. 
\end{proofof}

>>>>

\begin{proofof}[undecidable-non-negative:lem]
    Recall that by the undecidability of the Hilbert's Tenth Problem \cite[Problem 10 page 22]{HILB1902}, it is
    undecidable whether a given polynomial $P \in \Rel[\vec{X}]$ has a solution in
    $\Rel^n$ \cite{MATI1970,DAVIS1973}.
    The result follows by observing that
    checking whether a polynomial $P \in \Rel[\vec{X}]$ has no solutions in $\Rel^n$ if and only 
    if $P^2 - 1$ is \kl{non-negative}.

    Let us prove that checking whether $P \in \CoveredPoly$ is undecidable. Let
    $P \in \Rel[\vec{X}]$ be a polynomial, and let us write $P = P_+ - P_-$
    where both $P_+$ and $P_-$ belong to $\Nat[\vec{X}]$. Let $Z$ be a fresh
    indeterminate. We define $Q \defined P_+ (Z - 1) P_-$. By construction, $Q$
    has \kl{non-negative} \kl{maximal monomials}. Furthermore, $Q$ itself is
    \kl{non-negative} if and only if $\restr{Q}{Z = 0} = P$ is \kl{non-negative},
    since for every $n > 0$, $\restr{Q}{Z = n} \in \Nat[\vec{X}]$ is
    \kl{non-negative}. We conclude that 
    $P$ is \kl{non-negative} if and only if $Q$ is \kl{non-negative},
    if and only if $Q \in \CoveredPoly$.
\end{proofof}

### 2024-09-13 parts/correction.tex
Hilbert!

Let us briefly remark that \cref{corrected-version:thm} is surprising given the
fact that is not possible to decide whether a polynomial $P \in \Rel[\vec{X}]$
is \kl{non-negative} (\cref{undecidable-non-negative:lem}), by reduction to the
undecidability of Hilbert's Tenth Problem \cite{HILB1902,MATI1970}. That is,
$\CorrectPoly$ is a decidable class that interpolates between the undecidable
class of non-negative polynomials and strictly generalize polynomials in
$\Nat[\vec{X}]$.

\begin{remark}[restate=undecidable-non-negative:lem,label=undecidable-non-negative:lem]
    \proofref{undecidable-non-negative:lem}
    Checking whether a polynomial
    $P \in \Rel[\vec{X}]$ is \kl{non-negative} is undecidable.
\end{remark}

@@

>>>>

Let us briefly remark that \cref{corrected-version:thm} is surprising given the
fact that is not possible to decide whether a polynomial $P \in \Rel[\vec{X}]$
is \kl{non-negative} or if a polynomial $P$ belongs to $\CoveredPoly$
(\cref{undecidable-non-negative:lem}), by reduction to the undecidability of
Hilbert's Tenth Problem \cite{HILB1902,MATI1970}. That is, $\CorrectPoly$ is a
decidable class that strictly contains $\Nat[\vec{X}]$, and is contained in the
undecidable classes $\CoveredPoly$ and the class of \kl{non-negative}
polynomials.

\begin{remark}[restate=undecidable-non-negative:lem,label=undecidable-non-negative:lem]
    \proofref{undecidable-non-negative:lem}
    Checking whether a polynomial
    $P \in \Rel[\vec{X}]$ is \kl{non-negative} is undecidable.
    Similarly, checking whether a polynomial $P \in \Rel[\vec{X}]$
    belongs to $\CoveredPoly$ is undecidable.
\end{remark}

@@

We conclude from \cref{lem:correct-covered-2} that $\CoveredPoly[X,Y]$ is a
decidable class of polynomials, hence that Hilbert's Tenth Problem is
decidable for polynomials in two indeterminates with non-negative maximal
monomials.
